---
title: "Homework 7"
author: "Hannah Shafer"
date: "3/2/2022"
output: html_document
---
#HW 7 draft

1: examine iris data set:
```{r}
print(iris)
summary(iris)
```
150 observations (50 of each species) and 5 variables (sepal length, sepal width, petal length, petal width, species)

Load packages needed:
```{r}
library(tidyverse)
```

2: new dataframe with just virginica and versicolor, sepal lengths must be greater than 6, and sepal widths must be greater than 2.5
```{r}
iris1 <- filter(iris, Species %in% c("versicolor","virginica"), Sepal.Length > 6, Sepal.Width > 2.5)

print(iris1) 
summary(iris1)
```
56 observations and 5 variables

3: new dataframe with only species, sepal length, and sepal width variables
```{r}
iris2 <- select(iris1, Sepal.Length, Sepal.Width, Species)
print(iris2)
```
56 observations and 3 variables

4: order data frame by largest to smallest sepal length, print head
```{r}
iris3 <- arrange(iris2, by=desc(Sepal.Length))
head(iris3)
```

5: add new variable for sepal area
```{r}
iris4 <- mutate(iris3, Sepal.Area = Sepal.Length*Sepal.Width)
print(iris4)
```
56 observations and 4 variables

6: calculate avg sepal length, avg sepal width, and sample size of iris4
```{r}
iris5 <- summarize(iris4, meanSepalLength = mean(Sepal.Length), meanSepalWidth = mean(Sepal.Width), number=n())
print(iris5)
```

7: calculate avg sepal length, avg sepal width, and sample size by species
```{r}
iris6 <- iris4 %>%
  group_by(Species) %>%
  summarize(meanSepalLength=mean(Sepal.Length), meanSepalWidth=mean(Sepal.Width), number=n())
print(iris6)
```

8: use the pipe tool to get iris6 from iris1 with just one output
```{r}
filter(iris, Species %in% c("versicolor","virginica"), Sepal.Length > 6, Sepal.Width > 2.5) %>%
  select(Sepal.Length, Sepal.Width, Species) %>%
  arrange(by=desc(Sepal.Length)) %>%
  mutate(Sepal.Area = Sepal.Length*Sepal.Width) %>%
  group_by(Species) %>%
  summarize(meanSepalLength=mean(Sepal.Length), meanSepalWidge=mean(Sepal.Width), number=n())
```

9: longer data frame using iris but with three columns, Species, Measure, and Value
```{r}
iris %>%
  pivot_longer(cols=Sepal.Length:Petal.Width,
               names_to = "Measure",
               values_to = "Value")
```
