---
title: "Homework 10"
author: "Hannah Shafer"
date: "3/30/2022"
output: html_document
---

1. For loop within a function to find number of zeros in numeric vector
```{r}
####################################
# FUNCTION: zeros
# purpose: calculate the number of zeroes in a numeric vector
# input: x = numeric vector
# output: counter = number of 0s in vector
#----------------------------------
counter <- 0
vector <- c(0,1,2,3,0,4,5,6,0)

zeros <- function(x=vector) {
  for(i in seq_along(x)) {
    if (x[i] == 0) counter <- counter+1
  }
  return(counter)
}
zeros(vector)
```

2. Do 1 but with subsetting instead of function/loop
```{r}
count <- length(vector[vector==0])
print(count)
```

3. Function for matrix creation
```{r}
####################################
# FUNCTION: my_matrix
# purpose: create matrix populated with products of (row number) * (column number)
# input: i= rows; j= columns
# output: matrix
#----------------------------------
my_matrix <- function(i=10,
                      j=10) {
  mat <- matrix(nrow=i,ncol=j)
  for(i in 1:nrow(mat)) {
    for(j in 1:ncol(mat)) {
      mat[i,j] <- (i*j)
    }
  }
  return(mat)
}
my_matrix()
```

4.a. Simulate 3 groups of data with different means, place all data into a data frame
```{r}
group1 <- round(rnorm(n=10, mean=100, sd=5))
group2 <- round(rnorm(n=10, mean=200, sd=5))
group3 <- round(rnorm(n=10, mean=300, sd=5))
group <- rep(c("group1","group2", "group3"),each=10)
response <- c(group1,group2,group3)
d_frame <- data.frame(group, response)
print(d_frame)
```

4.b. Custom function to shuffle response variables and recalculate group means
```{r}
library(dplyr)
####################################
# FUNCTION: shuffle
# purpose: shuffle response variables, calculate new means
# input: data frame
# output: shuffled data frame & vector of means
#----------------------------------
shuffle_data <- function(z=d_frame) {
  shuffled <- data.frame(group,sample(z$response))
  means <- c(mean(z[,z$response == "group1"]),mean(z[,z$response == "group2"]),mean(z[,z$response == "group3"]))
  return(list(z,means))
}

# use dplyr
#shuffled <- shuffle_data(d_frame)

shuffle_data()

# c(mean(z[,z$response == "group1"]),mean(z[,z$response == "group2"]),mean(z[,z$response == "group3"]))

```

4.c. Repeat function in b 100 times and store new results in dataframe
```{r}
# for (i in 1:100) {
#   shuffle(d_frame)
# }
```

4.d. Create histogram to reflect data in c
```{r}

```

