---
title: "Homework 10"
author: "Hannah Shafer"
date: "3/30/2022"
output: html_document
---

1. For loop within a function to find number of zeros in numeric vector
```{r}
####################################
# FUNCTION: zeros
# purpose: calculate the number of zeroes in a numeric vector
# input: x = numeric vector
# output: counter = number of 0s in vector
#----------------------------------
counter <- 0
vector <- c(0,1,2,3,0,4,5,6,0)

zeros <- function(x=vector) {
  for(i in seq_along(x)) {
    if (x[i] == 0) counter <- counter+1
  }
  return(counter)
}
zeros(vector)
```

2. Do 1 but with subsetting instead of function/loop
```{r}
count <- length(vector[vector==0])
print(count)
```

3. Function for matrix creation
```{r}
####################################
# FUNCTION: my_matrix
# purpose: create matrix populated with products of (row number) * (column number)
# input: i= rows; j= columns
# output: matrix
#----------------------------------
my_matrix <- function(i=10,
                      j=10) {
  mat <- matrix(nrow=i,ncol=j)
  for(i in 1:nrow(mat)) {
    for(j in 1:ncol(mat)) {
      mat[i,j] <- (i*j)
    }
  }
  return(mat)
}
my_matrix()
```

4.a. Simulate 3 groups of data with different means, place all data into a data frame
```{r}
group1 <- round(rnorm(n=10, mean=100, sd=5))
group2 <- round(rnorm(n=10, mean=200, sd=5))
group3 <- round(rnorm(n=10, mean=300, sd=5))
group <- rep(c("group1","group2", "group3"),each=10)
response <- c(group1,group2,group3)
d_frame <- data.frame(group, response)
print(d_frame)
```

4.b. Custom function to shuffle response variables and recalculate group means
```{r}
library(dplyr)
####################################
# FUNCTION: shuffle
# purpose: shuffle response variables, calculate new means
# input: data frame
# output: shuffled data frame & vector of means
#----------------------------------
shuffle_data <- function(z=d_frame) {
  shuffled <- data.frame(group=z$group,respose=sample(z$response))
  means <- c(mean(shuffled[shuffled$group == "group1",2]),mean(shuffled[shuffled$group == "group2",2]),mean(shuffled[shuffled$group == "group3",2]))
  return(list(shuffled,means))
}
shuffle_data()
```

4.c. Repeat function in 4.b. 100 times and store new results in dataframe
```{r}
d.frame <- data.frame(iter=NA, mean1=NA, mean2=NA, mean3=NA)
for (i in 1:100) {
  shuffled <- shuffle_data(d_frame)
  unlist(shuffled[[2]])
  d.frame[i,] <- c(i, shuffled[[2]][1], shuffled[[2]][2], shuffled[[2]][3])
}
print(d.frame)
```

4.d. Create histograms to reflect data in c
```{r}
library(ggplot2)
qplot(x=d.frame$mean1,color=I("black"),fill=I("goldenrod"))
qplot(x=d.frame$mean2,color=I("black"),fill=I("goldenrod"))
qplot(x=d.frame$mean3,color=I("black"),fill=I("goldenrod"))

# Comparison to original means: from the graphs we can see that the new means are all centered around 200. This is same as the mean of group2, but far off from the mean of group1 (100) and the mean of group3 (300). This makes sense that it would center on the average of all the groups combined (mean of group1 + group2 + group3 = 200) - albeit with a larger average standard deviation - since all the responses were randomly shuffled.
```

